---
title: "Project"
author: "Maria Pais, Mónica Araújo"
date: "2023-11-07"
output:
  html_document:
    number_sections: yes
    theme: flatly  # journal flatly spacelab cosmo
    highlight: tango
    toc : yes
    toc_float:
      collapsed: true
      smoth_scroll: true
  word_document: 
    toc :yes
  pdf_document:
    toc : yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(readr)
library(dplyr)

# Load or create the loan_dev dataset
loan_dev <- read_delim("loan_dev.csv", delim = ";", 
    escape_double = FALSE, trim_ws = TRUE)
```

# Data Understanding eu

The data understanding phase of CRISP-DM involves taking a closer look at the data available for mining. This step is critical in avoiding unexpected problems during the next phase, data preparation, which is typically the longest part of a project.

Data understanding involves accessing the data and exploring it using tables and graphics. This enables you to determine the quality of the data and describe the results of these steps in the project documentation.

```{r cars}
summary(cars)
```

## Data Summarizing

The term Data Summarizing can be defined as the presentation of a summary/report of generated data in a comprehensible and informative manner.Data summarizing is typically numerical, visual or a combination of the two. It is a key skill in data analysis - we use it to provide insights both to others and to ourselves. 


Nesta linha de código vemos o número de empréstimos em que o contrato acabou sem problemas ("1") e aqueles em que o contrato acabou ou o empréstimo ainda não foi pago.

```{r loan_dev}
loan_dev %>%
     group_by(status) %>%
     count()
```

## Data Visualizing

Data visualization is the representation of data through use of common graphics, such as charts, plots, infographics, and even animations. These visual displays of information communicate complex data relationships and data-driven insights in a way that is easy to understand.

# Data Preparation monica

## Data quality and pre-processing

# Descriptive Modelling monica

# Predictive Modelling eu

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
